The Proofs

### 2023-11-13 proofs/main.tex
Finish the Algorithmic section
Restructure proofs

\usepackage[matha]{mathabx}

@@

\subsection{Equalities}
For simplicity, we assume alpha-equivalent terms equal.
This way, we assume that substitutions do not capture bound variables.
Besides, we equate
$[[∀pas.∀pbs.iN]]$ with $[[∀pas,pbs.iN]]$, 
as well as $[[∃nas.∃nbs.iP]]$ with $[[∃nas,nbs.iP]]$,
and lift these equations transitively and congruently 
to the whole system.

@@

$[[Γ ⊢ iP]]$ and $[[Γ ⊢ iN]]$ denote that the type is well-formed in the context $[[Γ]]$,
which, in fact, means that each free type variable of the type is contained in $[[Γ]]$
(it will be shown later in \cref{lemma:wf-soundness,lemma:wf-ctxt-equiv}).

@@

  \begin{multicols}{2}
  \ottdefnWFTNiWF{}
  \columnbreak

  \ottdefnWFTPiWF{}
  \end{multicols}

@@

\begin{definition}[Substitution]
  Substitutions (denoted as $[[σ]]$) 
  are represented by total functions form variables to types, preserving the polarity. 
\end{definition}

@@

\subsection{Declarative Subtyping}
Subtyping is one of the key mechanism of our system. 
It realizes the polymorphism: abstract $[[∀]]$ and 
$[[∃]]$ types can be used where concrete types are expected,
exactly because the subtyping relation between them.

@@

  \begin{multicols}{2}
    \ottdefnDOneNsub{}

    \ottdefnDOnePsup{}
  \end{multicols}
  \hfill

  \begin{multicols}{2}
    \ottdefnDOneNeq{}

    \ottdefnDOnePeq{}
  \end{multicols}
\end{definition}

@@

  \begin{multicols}{2}
  \ottdefnWFATNauWF{}
  \columnbreak

  \ottdefnWFATPauWF{}
  \end{multicols}

@@

  \begin{multicols}{2}
  \ottdefnEOneNeq{}
  \columnbreak\\
  \ottdefnEOnePeq{}
  \end{multicols}

@@

  \ottdefnONVar{}
  \ottdefnOPVar{}

@@

  \begin{multicols}{2}
  \ottdefnNrmNNorm{}
  \columnbreak\\
  \ottdefnNrmPNorm{}
  \end{multicols}

@@

  \begin{multicols}{2}
  \ottdefnNrmuNNorm{}
  \columnbreak\\
  \ottdefnNrmuPNorm{}
  \end{multicols}

@@

\begin{algorithm}[Subtyping]
  \label{alg:subtyping}
  \hfill\\
  \ottdefnANsub{}
  \ottdefnAPsup{}
\end{algorithm}

@@

\begin{definition}[Type Satisfying a Constraint Entry]
  \hfill\\
  \begin{multicols}{2}
  \ottdefnSATSCEN{}
  \columnbreak\\
  \ottdefnSATSCEP{}
  \end{multicols}
\end{definition}

@@

  \ottdefnSCME\\
\end{algorithm}

@@

\begin{algorithm}[Unification]
  \hfill
  \begin{multicols}{2}
  \ottdefnUNUnif{}
  \columnbreak\\
  \ottdefnUPUnif{}
  \end{multicols}
\end{algorithm}

@@

\begin{algorithm}[The Least Upper Bound Algorithm]
  \hfill\\
  \ottdefnLUBNsub{}
\end{algorithm}

@@

\begin{algorithm}[Type Upgrade]
  \hfill\\
  \ottdefnLUBUp{}
\end{algorithm}

@@

  \ottdefnAUAUP{}
  \ottdefnAUAUN{}

@@

\begin{definition}[Declarative Type Inference]
  \label{def:declarative-typing}
  \hfill
  \begin{multicols}{2}
  \ottdefnDTNInf{}
  \ottdefnDTPInf{}
  \ottdefnDTSpinInf{}
  \end{multicols}
\end{definition}

Let us discuss selected rules of the declarative system:
\begin{itemize}
  \item \ruleref{\ottdruleDTVarLabel}
    says that the type of a variable is inferred from the context.
  \item \ruleref{\ottdruleDTThunkLabel} says
    that the type of a thunk is inferred by shifting up the type of the 
    contained computation. Symmetrically, \ruleref{\ottdruleDTReturnLabel}
    infers the type of a return by shifting down the type of the
    contained value.
  \item \ruleref{\ottdruleDTPAnnotLabel} and \ruleref{\ottdruleDTNAnnotLabel} are symmetric.
    They allow the inferred type to be refined by annotating it with a supertype.
  \item \ruleref{\ottdruleDTNEquivLabel} and \ruleref{\ottdruleDTPEquivLabel}
    mean that th declarative system allows to infer any type from the equivalence class.
  \item \ruleref{\ottdruleDTUnpackLabel} is standard for existential types,
    and its first premise inferring the existential type of the value being unpacked.
    It is important however that the inferred existential type is normalized. 
    This is because there might be multiple equivalent existential types 
    with different order or even number of quantified variables, 
    and to bind them, the algorithm needs to fix the canonical one.
  \item \ruleref{\ottdruleDTAppLetAnnLabel} allows us to accommodate the applications
    with annotated let-bindings. The first premise infers the type of the head of the application,
    which must be a thunked computation. Then if after applying it 
    to the arguments, the resulting type can be equated to the annotated one,
    we infer the body of the let-binding in the context extended with the bound variable.
  \item \ruleref{\ottdruleDTAppLetLabel} is similar to \ruleref{\ottdruleDTAppLetAnnLabel},
    but it is used when the type of the application is unique, and thus, the annotation
    is redundant. Here $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ uniq]]$ means that
    if also $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ']]$ then $[[Γ ⊢ iQ ≈ iQ']]$
\end{itemize}

@@

\begin{algorithm}
  \hfill\\
  \ottdefnATNInf{}
  \ottdefnATPInf{}
  \ottdefnATSpinInf{}
\end{algorithm}

@@

Finally, let us discuss the algorithmic rules of the application inference:
\begin{itemize}
  \item \ruleref{\ottdruleATEmptyAppLabel}
    is the base case. If the list of arguments is empty, 
    the inferred type is the type of the head,
    and the algorithm returns it after normalizing.
  \item \ruleref{\ottdruleATArrowAppLabel}
    is the main rule of the algorithmic application inference.
    If the head has an arrow type $[[uQ → uN]]$,
    we find $[[SC1]]$---the minimal constraint ensuring that 
    $[[uQ]]$ is a supertype of the first argument's type.
    Then we make a recursive call applying $[[uN]]$ to the rest of the arguments,
    and merge the resulting constraint with $[[SC1]]$
  \item \ruleref{\ottdruleATForallAppLabel},
    analogously to the declarative case,
    is the rule ensuring the implicit elimination of the universal quantifiers. 
    This is the place where the algorithmic variables are generated.
    The algorithm simply replaces the quantified variables 
    $[[pas]]$ with fresh algorithmic variables $[[puas]]$,
    and makes a recursive call in the extended context. 
\end{itemize}

@@

\begin{algorithm}[Singular Constraint Entry]
  \hfill\\
  \ottdefnSINGscEP{}
  \ottdefnSINGscEN{}
\end{algorithm}

@@

@@

\subsection{Type Well-formedness}
\input{\genDir/tex/wf-lemmas.tex}

>>>>

% \usepackage[matha]{mathabx}

@@

\subsection{Equalities}
For simplicity, we assume alpha-equivalent terms are equal.
This way, we assume that substitutions do not capture bound variables.
Besides, we equate
$[[∀pas.∀pbs.iN]]$ with $[[∀pas,pbs.iN]]$, 
as well as $[[∃nas.∃nbs.iP]]$ with $[[∃nas,nbs.iP]]$,
and lift these equations transitively and congruently 
to the whole system.

@@

$[[Γ ⊢ iP]]$ and $[[Γ ⊢ iN]]$ denote that the type is well-formed in the context $[[Γ]]$,
which means that each free type variable of the type is contained in $[[Γ]]$
(it will be shown later in \cref{lemma:wf-soundness,lemma:wf-ctxt-equiv}).

@@

  \begin{multicols}{2}
  \ottdefnWFTNiWFLabeled{}
  \columnbreak

  \ottdefnWFTPiWFLabeled{}
  \end{multicols}

@@

\begin{definition}[Substitution]
  Substitutions (denoted as $[[σ]]$) 
  are represented by total functions from variables to types, preserving the polarity. 
\end{definition}

@@

\subsection{Declarative Subtyping}
Subtyping is one of the key mechanisms of our system. 
It realizes the polymorphism: abstract $[[∀]]$ and 
$[[∃]]$ types can be used where concrete types are expected,
exactly because of the subtyping relation between them.

@@

  \begin{multicols}{2}
    \ottdefnDOneNsubLabeled{}

    \ottdefnDOnePsupLabeled{}
  \end{multicols}
  \hfill

  \begin{multicols}{2}
    \ottdefnDOneNeqLabeled{}

    \ottdefnDOnePeqLabeled{}
  \end{multicols}
\end{definition}

@@

  \begin{multicols}{2}
  \ottdefnWFATNauWFLabeled{}
  \columnbreak

  \ottdefnWFATPauWFLabeled{}
  \end{multicols}

@@

  \begin{multicols}{2}
  \ottdefnEOneNeqLabeled{}
  \columnbreak\\
  \ottdefnEOnePeqLabeled{}
  \end{multicols}

@@

  \ottdefnONVarLabeled{}
  \ottdefnOPVarLabeled{}

@@

  \begin{multicols}{2}
  \ottdefnNrmNNormLabeled{}
  \columnbreak\\
  \ottdefnNrmPNormLabeled{}
  \end{multicols}

@@

  \begin{multicols}{2}
  \ottdefnNrmuNNormLabeled{}
  \columnbreak\\
  \ottdefnNrmuPNormLabeled{}
  \end{multicols}

@@

\begin{algorithm}[Subtyping]
  \label{alg:subtyping}
  \hfill\\
  \ottdefnANsubLabeled{}
  \ottdefnAPsupLabeled{}
\end{algorithm}

@@

\begin{definition}[Type Satisfying a Constraint Entry]
  \hfill\\
  \begin{multicols}{2}
  \ottdefnSATSCENLabeled{}
  \columnbreak\\
  \ottdefnSATSCEPLabeled{}
  \end{multicols}
\end{definition}

@@

  \ottdefnSCMELabeled\\
\end{algorithm}

@@

\begin{algorithm}[Unification]
  \hfill
  \begin{multicols}{2}
  \ottdefnUNUnifLabeled{}
  \columnbreak\\
  \ottdefnUPUnifLabeled{}
  \end{multicols}
\end{algorithm}

@@

\begin{algorithm}[The Least Upper Bound Algorithm]
  \hfill\\
  \ottdefnLUBNsubLabeled{}
\end{algorithm}

@@

\begin{algorithm}[Type Upgrade]
  \hfill\\
  \ottdefnLUBUpLabeled{}
\end{algorithm}

@@

  \ottdefnAUAUPLabeled{}
  \ottdefnAUAUNLabeled{}

@@

\begin{definition}[Declarative Type Inference]
  \label{def:declarative-typing}
  \ottdefnsDTLabeled
\end{definition}

Let us discuss the selected rules of the declarative system:
\begin{itemize}
  \item \ruleref{\ottdruleDTVarLabel}
    says that the type of a variable is inferred from the context.
  \item \ruleref{\ottdruleDTThunkLabel} says
    that the type of a thunk is inferred by shifting up the type of the 
    contained computation. Symmetrically, \ruleref{\ottdruleDTReturnLabel}
    infers the type of a return by shifting down the type of the
    contained value.
  \item \ruleref{\ottdruleDTPAnnotLabel} and \ruleref{\ottdruleDTNAnnotLabel} are symmetric.
    They allow the inferred type to be refined by annotating it with a supertype.
  \item \ruleref{\ottdruleDTNEquivLabel} and \ruleref{\ottdruleDTPEquivLabel}
    mean that the declarative system allows us to infer any type from the equivalence class.
  \item \ruleref{\ottdruleDTUnpackLabel} is standard for existential types,
    and its first premise infers the existential type of the value being unpacked.
    It is important however that the inferred existential type is normalized. 
    This is because there might be multiple equivalent existential types 
    with a different order or even number of quantified variables, 
    and to bind them, the algorithm needs to fix the canonical one.
  \item \ruleref{\ottdruleDTAppLetAnnLabel} allows us to accommodate the applications
    with annotated let-bindings. The first premise infers the type of the head of the application,
    which must be a thunked computation. Then if after applying it 
    to the arguments, the resulting type can be equated to the annotated one,
    we infer the body of the let-binding in the context extended with the bound variable.
  \item \ruleref{\ottdruleDTAppLetLabel} is similar to \ruleref{\ottdruleDTAppLetAnnLabel},
    but it is used when the type of the application is unique, and thus, the annotation
    is redundant. Here $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ uniq]]$ means that
    if also $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ']]$ then $[[Γ ⊢ iQ ≈ iQ']]$
\end{itemize}

@@

\begin{algorithm}
  \hfill\\
  \ottdefnATNInfLabeled{}
  \ottdefnATPInfLabeled{}
  \ottdefnATSpinInfLabeled{}
\end{algorithm}

@@

Finally, let us discuss the algorithmic rules of the application inference:
\begin{itemize}
  \item \ruleref{\ottdruleATEmptyAppLabel}
    is the base case. If the list of arguments is empty, 
    the inferred type is the type of the head,
    and the algorithm returns it after normalizing.
  \item \ruleref{\ottdruleATArrowAppLabel}
    is the main rule of algorithmic application inference.
    If the head has an arrow type $[[uQ → uN]]$,
    we find $[[SC1]]$---the minimal constraint ensuring that 
    $[[uQ]]$ is a supertype of the first argument's type.
    Then we make a recursive call applying $[[uN]]$ to the rest of the arguments,
    and merge the resulting constraint with $[[SC1]]$
  \item \ruleref{\ottdruleATForallAppLabel},
    analogously to the declarative case,
    is the rule ensuring the implicit elimination of the universal quantifiers. 
    This is the place where the algorithmic variables are generated.
    The algorithm simply replaces the quantified variables 
    $[[pas]]$ with fresh algorithmic variables $[[puas]]$,
    and makes a recursive call in the extended context. 
\end{itemize}

@@

\begin{algorithm}[Singular Constraint Entry]
  \hfill\\
  \ottdefnSINGscEPLabeled{}
  \ottdefnSINGscENLabeled{}
\end{algorithm}

@@

\section{Theorem Statements: Declarative}
\subsection{Type Well-Formedness}
\input{\genDir/tex/wf-statements.tex}

\subsection{Substitution}
\input{\genDir/tex/subst-statements.tex}

\subsection{Declarative Subtyping}
\input{\genDir/tex/decl-subtyping-statements.tex}

\subsection{Equivalence}
\label{sec:decl-equiv-statements}
\input{\genDir/tex/equiv-statements.tex}

\subsection{Variable Ordering}
\input{\genDir/tex/ord-statements.tex}

\subsection{Normaliztaion}
\input{\genDir/tex/norm-statements.tex}

\section{Declarative Typing}
\input{\genDir/tex/decl-typing-statements.tex}

\section{Theorem Statements: Algorithmic}

\subsection{Algorithmic Type Well-formedness}
\input{\genDir/tex/wf-algo-statements.tex}

\subsection{Algorithmic Substitution}
\input{\genDir/tex/subst-algo-statements.tex}

\subsection{Algorithmic Normalization}
\input{\genDir/tex/norm-alg-statements.tex}

\subsection{Algorithmic Equivalence}
\input{\genDir/tex/equiv-alg-statements.tex}

\subsection{Unification Constraint Merge}
\input{\genDir/tex/unif-constraint-merge-statements.tex}

\subsection{Unification}
\input{\genDir/tex/unification-statements.tex}

\subsection{Anti-unification}
\input{\genDir/tex/au-statements.tex}

\subsection{Upper Bounds}
\label{sec:alg-upper-bounds-statements}
\input{\genDir/tex/lub-statements.tex}

\subsection{Upgrade}
\label{sec:upgrade-statements}
\input{\genDir/tex/upgrade-statements.tex}

\subsection{Constraint Satisfaction}
\input{\genDir/tex/constraint-sat-statements.tex}

\subsection{Positive Subtyping}
\input{\genDir/tex/alg-pos-subtyping-statements.tex}

\subsection{Subtyping Constraint Merge}
\input{\genDir/tex/constraint-merge-statements.tex}

\subsection{Negative Subtyping}
\input{\genDir/tex/alg-neg-subtyping-statements.tex}

\subsection{Singularity}
\input{\genDir/tex/singularity-statements.tex}

\subsection{Correctness of the Typing Algorithm}
\input{\genDir/tex/alg-typing-statements.tex}

@@

\subsection{Type Well-Formedness}
\input{\genDir/tex/wf-lemmas.tex}

### 2023-11-13 proofs/tex/alg-neg-subtyping-lemmas.tex
Restructure proofs

\begin{observation}[Negative Algorithmic Subtyping is Deterministic]
    \label{obs:neg-subt-deterministic}
    For fixed $[[Γ]]$, $[[Θ]]$, $[[iM]]$, and $[[uN]]$, 
    if $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$ 
    and $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC']]$
    then $[[SC]]$ = $[[SC']]$.
\end{observation}
\begin{proof}
    First, notice that the shape of
    the input uniquely determines the rule applied to infer
    $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$,
    which is consequently, the same rule used to
    infer $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC']]$.

@@

\begin{lemma}[Soundness of Negative Subtyping] \label{lemma:neg-subt-soundness}
        If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iM]]$, $[[Γ ; dom(Θ) ⊢ uN]]$ and 
        $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$, then 
        $[[Θ ⊢ SC : uv(uN)]]$ and 
        for any normalized $[[uσ]]$ such that $[[ Θ  ⊢ uσ : SC ]]$,
        $[[ Γ ⊢ [uσ]uN ≤ iM ]]$.
\end{lemma}
\begin{proof}
    We prove it by induction on $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$.

@@

\begin{lemma}[Completeness of the Negative Subtyping]
    \label{lemma:neg-subt-completeness}
    Suppose that $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iM]]$, $[[Γ ; dom(Θ) ⊢ uN]]$,
    and $[[uN]]$ does not contain negative unification variables ($[[α̂⁻]] \notin [[uv uN]]$).
    Then for any $[[Θ ⊢ uσ : uv(uN)]]$ such that $[[Γ ⊢ [uσ]uN ≤ iM]]$,
    there exists $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$ and moreover, $[[ Θ ⊢ uσ : SC ]]$.
\end{lemma}
\begin{proof}
    We prove it by induction on $[[ Γ ⊢ [uσ]uN ≤ iM ]]$.
    Let us consider the last rule used in the derivation of $[[ Γ ⊢ [uσ]uN ≤ iM ]]$.
    \begin{caseof}
        \item $[[ Γ ⊢ [uσ]uN ≤ iM ]]$ is derived by \ruleref{\ottdruleDOneShiftULabel}\\
            Then $[[ uN ]] = [[ ↑uP ]]$, since
            the substitution $[[ [uσ]uN ]]$ must preserve the 
            top-level constructor of $[[uN]] \neq [[α̂⁻]]$ (since by assumption, $[[α̂⁻]] \notin [[uv uN]]$), 
            and $[[uQ]] = [[ ↓iM ]]$, and by inversion, $[[ Γ ⊢ [uσ]uN ≈ iM ]]$.
            The rest of the proof is symmetric to \cref{case:pos-subt-complete-upshift} of
            \cref{lemma:pos-subt-completeness}: notice that the algorithm does not make a recursive call, 
            and the difference in the induction statement for the positive and 
            the negative case here does not matter.

>>>>

\obsNegSubtDeterministic*
\begin{proof}
    First, notice that the shape of
    the input uniquely determines the rule applied to infer
    $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$,
    which is consequently, the same rule used to
    infer $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC']]$.

@@

\lemNegSubtSoundness*
\begin{proof}
    We prove it by induction on $[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$.

@@

\lemNegSubtCompleteness*
\begin{proof}
    We prove it by induction on $[[ Γ ⊢ [uσ]uN ≤ iM ]]$.
    Let us consider the last rule used in the derivation of $[[ Γ ⊢ [uσ]uN ≤ iM ]]$.
    \begin{caseof}
        \item $[[ Γ ⊢ [uσ]uN ≤ iM ]]$ is derived by \ruleref{\ottdruleDOneShiftULabel}\\
            Then $[[ uN ]] = [[ ↑uP ]]$, since
            the substitution $[[ [uσ]uN ]]$ must preserve the 
            top-level constructor of $[[uN]] \neq [[α̂⁻]]$ (since by assumption, $[[α̂⁻]] \notin [[uv uN]]$), 
            and $[[uQ]] = [[ ↓iM ]]$, and by inversion, $[[ Γ ⊢ [uσ]uN ≈ iM ]]$.
            The rest of the proof is symmetric to \cref{case:pos-subt-complete-upshift} of
            \cref{lemma:pos-subt-completeness}: notice that the algorithm does not make a recursive call, 
            and the difference in the induction statement for the positive and 
            the negative case here does not matter.

### 2023-11-13 proofs/tex/alg-pos-subtyping-lemmas.tex
Restructure proofs

\begin{observation}[Positive Subtyping is Deterministic]
    \label{obs:pos-subt-deterministic}
    For fixed $[[Γ]]$, $[[Θ]]$, $[[uP]]$, and $[[iQ]]$,
    if $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ and $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC']]$
    then $[[SC]] = [[SC']]$.
\end{observation}
\begin{proof}
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$.
    First, it is easy to see that the rule applied to infer 
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ uniquely depends on the input, 
    and those, it is the same rule that is inferring
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC']]$.
    Second, the premises of each rule are deterministic on the input: 
    unification is deterministic by \cref{obs:unif-deterministic},
    upgrade is deterministic by \cref{obs:upgrade-deterministic},
    the choice of the fresh algorithmic variables is deterministic by
    convention, as discussed in \cref{sec:fresh-selection},
    positive subtyping by the induction hypothesis.
\end{proof}

\begin{lemma}[Soundness of the Positive Subtyping] \label{lemma:pos-subt-soundness}
    If $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$, $[[Γ ; dom(Θ) ⊢  uP]]$, and 
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$,
    then $[[Θ ⊢ SC : uv uP]]$ and
    for any normalized $[[uσ]]$ such that $[[ Θ ⊢ uσ : SC ]]$,
    $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
\end{lemma}
\begin{proof} 
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$. 
    Let us consider the last rule to infer this judgment.
    \begin{caseof}
    \item \ruleref{\ottdruleAPUVarLabel} then
        $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ â⁺ ≥ iP' ⫤ (â⁺ :≥ iQ')]]$ where
        $[[â⁺[Δ] ∊ Θ]]$ and $[[upgrade G ⊢ iP' to Δ = iQ']]$.

@@

\begin{lemma}[Completeness of the Positive Subtyping] \label{lemma:pos-subt-completeness}
    Suppose that $[[Γ ⊢ Θ]]$, $[[Γ ⊢ iQ]]$ and $[[Γ ; dom(Θ) ⊢  uP]]$.
    Then for any $[[Θ ⊢ uσ : uv(uP)]]$ such that $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$,
    there exists $[[Γ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ and moreover, $[[ Θ ⊢ uσ : SC ]]$.
\end{lemma}
\begin{proof}
    Let us prove this lemma by induction on $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
    Let us consider the last rule used in the derivation,
    but first, consider the base case for the substitution $[[ [uσ]uP ]]$:
    \begin{caseof}
        \item \label{case:pos-subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ 
            (for potentially empty $[[nbs]]$)\\
            Then by assumption, $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ iQ ]]$ (where $[[ {nbs} ∩ fv [uσ]α̂⁺ = ∅]]$).
            Let us decompose $[[iQ]]$ as $[[iQ]] = [[∃ncs.iQ0]]$, where $[[iQ0]]$ does
            not start with $[[∃]]$.

>>>>

\obsPosSubtDeterministic*
\begin{proof}
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$.
    First, it is easy to see that the rule applied to infer 
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ uniquely depends on the input, 
    and those, it is the same rule that is inferring
    $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC']]$.
    Second, the premises of each rule are deterministic on the input: 
    unification is deterministic by \cref{obs:unif-deterministic},
    upgrade is deterministic by \cref{obs:upgrade-deterministic},
    the choice of the fresh algorithmic variables is deterministic by
    convention, as discussed in \cref{sec:fresh-selection},
    positive subtyping by the induction hypothesis.
\end{proof}

\lemPosSubtSoundness*
\begin{proof} 
    We prove it by induction on $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$. 
    Let us consider the last rule to infer this judgment.
    \begin{caseof}
    \item \ruleref{\ottdruleAPUVarLabel} then
        $[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$ has shape $[[Γ;Θ ⊨ â⁺ ≥ iP' ⫤ (â⁺ :≥ iQ')]]$ where
        $[[â⁺[Δ] ∊ Θ]]$ and $[[upgrade G ⊢ iP' to Δ = iQ']]$.

@@

\lemPosSubtCompleteness*
\begin{proof}
    Let us prove this lemma by induction on $[[ Γ ⊢ [uσ]uP ≥ iQ ]]$.
    Let us consider the last rule used in the derivation,
    but first, consider the base case for the substitution $[[ [uσ]uP ]]$:
    \begin{caseof}
        \item \label{case:pos-subt-complete-base} $[[uP]] = [[ ∃nbs.α̂⁺ ]]$ 
            (for potentially empty $[[nbs]]$)\\
            Then by assumption, $[[ Γ ⊢ ∃nbs.[uσ]α̂⁺ ≥ iQ ]]$ (where $[[ {nbs} ∩ fv [uσ]α̂⁺ = ∅]]$).
            Let us decompose $[[iQ]]$ as $[[iQ]] = [[∃ncs.iQ0]]$, where $[[iQ0]]$ does
            not start with $[[∃]]$.
